---
title: "p8105_hw6_cac2225"
author: "Courtney Chan"
date: "November 26, 2018"
output: github_document
---

#Homework 6

##Problem 0
A github repo and local R project was created, an Rmd file was produced, knitted, and committed to github.

##Problem 1

First loading all necessary packages for the homework assignment.

```{r loading necessary packages}
library(tidyverse)
library(modelr)
library(mgcv)

```

The csv file is imported using the read_csv function and the dataset is cleaned.

```{r}

homicide = read_csv("./data/homicide-data.csv") %>% 
  janitor::clean_names()

```


A city_state variable is created using the functions paste and mutate. Also a "solved" variable is created to indicate which cases have or have not been solved based on homework 5's prompt. The number of unsolved homicides is defined as those for which the disposition is “Closed without arrest” or “Open/No arrest”). City_states Dallas, TX, Pheonix, AZ, Kansas City, MO and Tulsa, AL are filtered out of the dataset.


```{r creating city_state variable, mutating solved, filtering city_state 2}

homicide = homicide %>% 
  mutate(city_state = str_c(city,", ",state)) %>% 
  mutate(solved = recode(disposition, "Closed without arrest" = "Unsolved", "Open/No arrest" = "Unsolved", "Closed by arrest" = "Solved")) %>% 
  filter(city_state != "Dallas, TX", city_state != "Phoenix, AZ", city_state != "Kansas City, MO", city_state != "Tulsa, AL")
  
```
The victim race variable is modified to be binary and for the catagory "White" to be defined as the reference catagory. Hispanic, Asian, Black, Other and Unknown were coded as "Non-white".

```{r modifying victim_race variable }

homicide = homicide %>% 
  mutate(victim_race = recode(victim_race, "Black" = "Non-white", "Hispanic" = "Non-white", "Asian" = "Non-white", "Other" = "Non-white", "Unknown" = "Non-white")) %>%  mutate(victim_race = factor(victim_race, levels = c("Non-white", "White")))

is.factor(homicide$victim_race)

```
Victim_age is first checked to see whether it is a numeric variable. Since it is not, the "unknown" category is first recoded as "NA". Then victim_age is defined as a numeric variable. It is checked again to see whether it was converted successfully into an numeric variable.

```{r victim_age}

is.numeric(homicide$victim_age)

homicide = homicide %>% 
 mutate(victim_age = as.numeric(victim_age)) %>% 
  mutate(factor(solved)) %>% 
  mutate(solved = factor(solved, levels = c("Solved", "Unsolved")))

is.numeric(homicide$victim_age)

is.factor(homicide$solved)

```

```{r baltimore analysis alone}

baltimore = homicide %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(solved ~ victim_age +  victim_sex + victim_race, data = ., family = binomial())

baltimore %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>% 
  mutate(conf.low = exp(estimate - (1.96*std.error)),
         conf.high = exp(estimate + (1.96*std.error))) %>% select(term, estimate, OR, conf.low, conf.high) %>% 
  filter(term == "victim_raceWhite")

```

For cases in Baltimore, MD, the point estimate describing the relationship between the outcome of solved homicides and the predictor victim race, adjusting for victim age and victim sex is 0.441, with a 95% confidence interval of 0.313 to 0.62.

For cases in Baltimate, MD, the odds of having a solved case amongst those who are non-white is 0.441 times the odds of having a solved case amongst those who are white, adjusting for victim age and sex. We are 95% confident that the true odds ratio lies between 0.313 and 0.62.

```{r applying glm to selected homicide df}

df_homicide = 
  homicide %>%
  group_by(city_state) %>% 
  nest() %>% 
  mutate(model = map(data, ~glm(solved ~ victim_age +  victim_sex + victim_race, data = ., family = binomial())), 
model = map(model, broom::tidy)) %>% 
  select(-data) %>% 
  unnest() %>% 
   mutate(OR = exp(estimate)) %>% 
  mutate(conf.low = exp(estimate - (1.96*std.error)),
         conf.high = exp(estimate + (1.96*std.error))) %>% select(city_state, term, estimate, OR, conf.low, conf.high) %>% 
  filter(term == "victim_raceWhite")


```

Glm is applied to each city_state and an OR with 95% confidence intervals is calculated for each city_state, comparing the odds of the case being solved, between non-whites and whites, adjusting for sex and age. 

```{r plot estimate ORs and CIs for each city}

df_homicide %>% 
  ggplot(aes(fct_reorder(city_state, desc(OR)), OR)) + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + theme(axis.text.x = element_text(angle = 90)) + ggtitle("estimated adjusted ORs and confidence intervals per city, state") + xlab("city, state") + ylab("adjusted estimate ORs of solved cases, non- vs whites")

```

##Problem 2

###Uploading the datafile, and dataset preparation/cleaning

The csv file is imported using the read_csv function and the dataset is cleaned.

```{r importing data}

birthweight = read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names()

```

```{r looking at the variable type}
birthweight

```

According to the output, the variables are all integers. Looking at the data dictionary, variables that should be converted into factors should be baby sex, father's race, presence of malformations and mother's race.

```{r recoding variables and converting variables into factors}

birthweight = birthweight %>% 
  mutate(babysex = recode(babysex, "1" = "male", "2" = "female"),
         frace = recode(frace, "1" = "white", "2" = "black", "3" = "asian", "4" = "puerto rican", "8" = "other", "9" = "unknown"),
         mrace = recode(mrace, "1" = "white", "2" = "black", "3" = "asian", "4" = "puerto rican", "8" = "other"),
         malform = recode(malform, "0" = "absent", "1" = "present"))


birthweight = birthweight %>% 
  mutate(babysex = factor(babysex, levels = c("female", "male"))) %>% 
           mutate(malform = factor(malform, levels = c("present", "absent")))
         
birthweight = birthweight %>% 
         mutate(frace = factor(frace))

birthweight = birthweight %>% 
         mutate(mrace = factor(mrace))

is.factor(birthweight$babysex)
is.factor(birthweight$malform)
is.factor(birthweight$frace)
is.factor(birthweight$mrace)

```

Checking for missing data within the birthweight dataset.

```{r checking for missing data}

skimr::skim(birthweight)

```

Looking at the missing column for each variable, there are no missing values.

###Proposed model

There is some literature evidence that certain variables are associated with birthweight of infants. Maternal smoking is known to contribute to low birthweight thus in this model it would be important to include this variable. Maternal pre-pregnancy BMI may impact birthweight, with being underweight during pre-pregnancy leading to low birthweight babies and being overweight during pre-pregnancy leading to large for gestational age babies. Amount of gestational weight gained also impacts birthweight. However, it would be interesting to determine which variable has a stronger impact on birthweight.

This proposed model should thus include the following variables, to determine their association with birthweight: maternal pre-pregnancy BMI, gestational weight gain and maternal smoking during pregnancy.

```{r initial model birthweight}

lm_birthweight = lm(bwt ~  wtgain + ppbmi + smoken, data = birthweight)

```

To determine whether linear assumptions apply to this model, the residuals and fitted values are calculated and plotted.

First the residual values are plotted against each predictor variable. The residual values for wtgain, ppbmi and smoken all appear to have non-random patterns. This proposed linear model may not be the best fit for this set of data. The residuals for the wtgain plot clearly clusters towards the middle of the graph. The residuals for ppbmi and smoken cluster towards the left of their respective plots.


```{r wtgain vs residuals}

birthweight %>% 
  modelr::add_residuals(lm_birthweight) %>% 
  ggplot(aes(x = wtgain, y = resid)) + geom_point() +   labs(
    x = "predictor weight gain",
    y = "residuals",
    title = "scatterplot of weight gain variable vs residuals")

```

```{r ppbmi vs residuals}

birthweight %>% 
  modelr::add_residuals(lm_birthweight) %>% 
  ggplot(aes(x = ppbmi, y = resid)) + geom_point() + labs(
    x = "predictor pre pregnancy bmi",
    y = "residuals",
    title = "scatterplot of pre pregnancy bmi vs. residuals")

```

```{r smoken vs residuals}

birthweight %>% 
  modelr::add_residuals(lm_birthweight) %>% 
  ggplot(aes(x = smoken, y = resid)) + geom_point() + labs(
    x = "predictor smoken",
    y = "residuals",
    title = "scatterplot of avg number of cigarettes per day during pregnancy vs residuals")

```

```{r overall distribution of residuals}

birthweight %>% 
  modelr::add_residuals(lm_birthweight) %>% 
  ggplot(aes(x = resid)) + geom_density() + labs(
    x = "residuals",
    y = "density",
    title = "density plot of residuals")

```

Looking at the density plot of the residuals alone, it seems although they center around 0, there is a slight skew to the left.

Additionally, the residuals are plotted against the fitted values.

```{r fitted values vs. residuals}

birthweight %>% 
  modelr::add_predictions(lm_birthweight) %>% 
  modelr::add_residuals(lm_birthweight) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point() + labs(
    x = "fitted values",
    y = "residuals",
    title = "scatterplot of fitted values vs residuals")

```

The residuals do not appear randomly scattered around 0, in fact there seems to be some clustering slightly towards the left of the plot.

This proposed model therefore is definitely not perfect. Non-linear models should be tested to see if they fit the data better. Each variable should be examined for outliers, which should be removed and the model rerun.


###Comparison of models

Creating and running the two other comparison linear models.

```{r two other linear models for comparison}

lm_one = lm(bwt ~  blength + gaweeks, data = birthweight)

lm_two = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight)

```

The cross validation function is used to compare the three models and rsme values are calculated to quantify the fit of each model.

```{r cross validation of the three models}
cv_bwt = 
  crossv_mc(birthweight, 100)

cv_bwt = 
  cv_bwt %>% 
  mutate(lm_birthweight = map(train, ~lm(bwt ~  wtgain + ppbmi + smoken, data = .x)),
         lm_one = map(train, ~lm(bwt ~  blength + gaweeks, data = .x)),
         lm_two = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = .x))) %>% 
  mutate(rmse_bwt = map2_dbl(lm_birthweight, test, ~rmse(model = .x, data = .y))) %>% 
           mutate(rmse_one = map2_dbl(lm_one, test, ~rmse(model = .x, data = .y))) %>% 
           mutate(rmse_two = map2_dbl(lm_two, test, ~rmse(model = .x, data = .y)))

```

```{r violin plots of rsme values for each model}

cv_bwt %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() + labs(
    title = "model versus rmse value")

```

Looking at the violin plots of the rmses, two has the lowest rmse value thus fits this data the best. My model's rmse is ridiculously high which confirms my previous diagnostic steps and thus needs to be seriously re-evaluated.
